\documentclass[12pt, a4paper]{article}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[margin=2.5cm]{geometry}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{cleveref}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black,
}

\onehalfspacing

% --- Title ---
\title{Patents, Papers, Parts \& Planet:\\
  Tracing Synthetic Biology Innovation Trajectories\\
  from Student Projects to Patents}

\author{[Author Name]\\
  \small CSH Vienna / TU Straubing\\
  \small \href{mailto:your.email@example.com}{your.email@example.com}}

\date{\today}

% ============================================================
\begin{document}
% ============================================================

\maketitle

\begin{abstract}
  This paper studies whether student iGEM projects, academic publications,
  and patents within synthetic biology form semantically related local
  innovation trajectories at the city level. We construct a corpus of
  papers (OpenAlex), patents (Lens.org), and student projects (iGEM Registry),
  encode them with a shared sentence-transformer embedding, and cluster the
  joint embedding space. We ask whether the three artifact types occupy
  related regions of semantic space, share thematic clusters, and show
  plausible city-level temporal continuity. Carbon capture in synthetic
  biology serves as the primary case study. We find \ldots\ [to be completed].
  These findings suggest \ldots\ [to be completed].
\end{abstract}

\textbf{Keywords:} synthetic biology, innovation trajectories, knowledge geography,
text embeddings, iGEM, patent analysis, carbon capture

\tableofcontents
\newpage

% ============================================================
\section{Introduction}
\label{sec:intro}
% ============================================================

Innovation in frontier technologies rarely happens in isolation.
It tends to cluster geographically, to follow temporal sequences,
and to leave traces across multiple types of knowledge artifact:
academic papers, patents, and — more recently — open-science platforms
like the iGEM (International Genetically Engineered Machine) student
competition \citep{igem}.

This paper asks a specific question: within synthetic biology,
do student projects, academic publications, and patents form
\emph{semantically related local innovation trajectories}?
The working hypothesis is that innovation in synthetic biology
may follow a local path in which:
\begin{enumerate}
  \item student projects introduce early ideas,
  \item academic publications develop and refine them,
  \item and patents reflect downstream application or commercial translation.
\end{enumerate}

We do not claim to establish causality.
We use language of \emph{semantic relatedness}, \emph{association},
\emph{cluster overlap}, and \emph{plausible temporal sequence}.

The primary case study is \textbf{carbon capture in synthetic biology}:
the engineering of microorganisms to fix CO\textsubscript{2} or produce
carbon-neutral fuels. This subfield is large enough to find examples
across all three artifact types and is directly relevant to planetary
sustainability.

The analysis combines three data sources:
(1) papers from OpenAlex \citep{priem2022openalex},
(2) patents from Lens.org,
and (3) iGEM project and parts data from the iGEM Registry.
All artifacts are encoded with a shared sentence-transformer embedding
\citep{reimers2019sentencebert} and projected to a shared semantic space
using UMAP \citep{mcinnes2018umap}.

% ============================================================
\section{Background and Related Work}
\label{sec:background}
% ============================================================

\subsection{A Field That Named Itself in a Hurry}

In the year 2000, two groups of researchers published papers in the same
issue of \textit{Nature} describing things that did not previously exist:
a genetic toggle switch and a synthetic oscillator, both running inside
living \textit{E.~coli} cells \citep{gardner2000toggle,elowitz2000synthetic}.
The universe, which does not typically trouble itself with editorial timing,
had apparently decided to make a point.
What both papers demonstrated was not merely that you could manipulate the
genes of a bacterium---this had been possible since the 1970s---but that
you could redesign the \textit{regulatory logic} of a living cell as one
redesigns a circuit board.
Not modify the logic.
Not nudge it.
Redesign it, from specifications, on purpose.

The field that crystallised around these demonstrations acquired its current
name somewhat opportunistically.
``Synthetic biology'' had appeared before---the phrase carries prior
incarnations in the early twentieth century and again in the 1970s, when it
described directed genetic modification more generically
\citep{cameron2014brief}.
What followed 2000 was different in ambition: an explicit engineering
programme.
Where earlier biology had described life and molecular biology had unpicked
its components, synthetic biology announced the intention to \textit{design}
it from first principles, following the same logic used to design integrated
circuits or suspension bridges.
This was, by any reasonable assessment, an extraordinary thing to announce,
and it was announced with corresponding enthusiasm.

\citet{endy2005foundations} articulated the engineering agenda most clearly:
abstraction hierarchies, standardised parts, decoupled design and manufacture.
The BioBricks parts standard followed, defining a format in which genetic
components could be combined and recombined across projects---a metaphor
involving Lego that was irresistible even then, and has since been thoroughly
exhausted by press releases.
In 2003, an MIT course gave rise to iGEM, a student competition that would
grow from five teams to several hundred within a decade, producing an open
repository of biological parts and demonstrating that undergraduate students
with a fermentation flask and a standard parts registry could contribute
genuine designs to a frontier research field \citep{igem}.

It is worth pausing on this last point.
iGEM's artefacts are publicly deposited, documented, and traceable.
A team in Chiba in 2005 could build on a BioBrick registered by a team in
Edinburgh in 2004.
The Registry of Standard Biological Parts is, in this respect, a kind of
distributed laboratory notebook for an entire generation of apprentice
engineers.
Whether those apprentices \textit{lead} subsequent research, or
\textit{follow} it, or simply work in parallel with it, is one of the
questions this paper exists to ask.

\subsection{The Geography of Clever Ideas}

If synthetic biology is unusual for being an engineering discipline grown
inside a biological science, it is, in this respect, a perfectly ordinary
instance of a phenomenon that economists of innovation have studied for over
a century: the tendency of knowledge-intensive activity to cluster in space.

\citet{marshall1890principles} identified the phenomenon in Victorian
industrial districts and could not quite explain it.
His successors fared better.
\citet{jaffe1993geographic} showed that patent citations cluster
geographically: patents are significantly more likely to cite other patents
from the same metropolitan area than the local share of knowledge production
would predict.
The implication is uncomfortable but hard to avoid.
Knowledge, despite being ostensibly weightless, travels poorly across
distance.
It sticks to the coffee machines, the seminar rooms, the chance conversations
between postdocs on the way to the cafeteria.
\citet{audretsch1996r} confirmed that R\&D spillovers attenuate with
distance across industries broadly, not only in any single sector.

The mechanism is imperfectly understood but generally attributed to tacit
knowledge: the kind of expertise that cannot be written down, that lives in
the hands and intuitions of practitioners, and that is therefore transmitted
face-to-face or not at all \citep{polanyi1966tacit}.
A paper or a patent is an artefact---a crystallised residue of what was once
tacit knowledge and has been partially made explicit.
A city is where the tacit knowledge lives that produced the artefact.

\citet{bathelt2004clusters} described this ambient propagation as
``local buzz'': the continuous circulation of information within a geographic
cluster of related firms and institutions, enabling rapid diffusion of
knowledge, norms, and research agendas that would travel far more slowly
between cities.
The implication for a field like synthetic biology is that innovation
trajectories---the sequence of ideas from early exploration to commercial
application---should show spatial coherence.
Artefacts from the same city should concern related problems more often than
would be expected by chance, because the people who produced them were
embedded in the same local knowledge environment.
Whether this prediction holds for three very different artefact types---student
projects, academic publications, and patents---is the empirical question at
the heart of the present analysis.

\subsection{The Peculiar Case of the Student Engineer}

iGEM occupies an anomalous position in the innovation system.
It is not a conventional university research programme, not an industry
R\&D pipeline, and not an outreach exercise, though it has elements of all
three.
Its participants are undergraduates and postgraduates doing original
biological engineering on an annual cycle, with results deposited in a public
registry and judged at a global jamboree.
Viewed with sufficient distance, the organisational form is rather odd:
a kind of open-source hardware movement for living systems, in which the
barrier to entry is a functioning university and a willing supervisor.

The canonical linear account of innovation places basic research upstream
and application downstream, with knowledge flowing from academic publication
to commercial patent.
The iGEM model complicates this picture in at least two ways.
First, iGEM projects are designed rather than discovered: teams identify an
application and engineer the biology to enable it.
This is closer to the logic of downstream application than upstream
exploration---yet the participants are students, not commercial engineers.
Second, iGEM teams routinely work with cutting-edge capabilities; a team
implementing a CRISPR-based diagnostic in 2013 or a carbon-fixation pathway
in 2010 was working at the research frontier, not in a training room.

What we do not have, at present, is a systematic characterisation of whether
iGEM project themes correspond to the themes of concurrent or subsequent
academic publication and commercial patenting---and in particular whether
that correspondence, where it exists, is geographically structured.
This is a gap that the present analysis proposes to address: modestly,
computationally, and with the appropriate epistemic hedges firmly in place.

\subsection{Measuring a Field That Keeps Moving}

Finding ``all the synthetic biology'' in a large archive turns out to be a
methodological problem of considerable subtlety.
The field's terminological promiscuity is part of the difficulty.
Synthetic biology shares vocabulary with genetic engineering, metabolic
engineering, molecular biology, and protein design to an extent that makes
keyword search without careful scoping an exercise in catching everything
while meaning to catch something specific.

\citet{shapira2017tracking} addressed this directly by distinguishing core
self-identifying terms---``synthetic biology'', ``synthetic genomics''---from
subfield terms that preceded or transcend the label: ``BioBrick'',
``repressilator'', ``minimal genome''.
Broad terms like ``metabolic engineering'' are excluded from retrieval; they
would import entire adjacent literatures with only partial overlap.
This two-layer approach underlies the paper corpus construction used here.

The patent literature presents a distinct but related problem.
Patent language is strategically expansive---applicants are motivated to
claim the widest possible scope---and synthetic biology patents in particular
overlap heavily with general biotechnology and genetic engineering.
\citet{oldham2018synbio} found that pure keyword searches radically
overestimate synthetic biology patent activity because the vocabulary is shared
across disciplines.
\citet{vandoren2013synbio}, taking a more structured approach, combined
keyword groups with an International Patent Classification (IPC) scope filter
restricted to biologically relevant class codes (C12N, C12P, C12Q, C12S,
C40B), using AND logic.
The intersection of ``resembles synthetic biology by keywords'' and ``is
classified in the correct technical domain'' is substantially more precise
than either criterion alone.
The patent corpus construction used here follows \citeauthor{vandoren2013synbio}
directly.

These methodological choices are not technical decoration.
They determine what the field \textit{is}, operationally, for the purpose of
everything that follows.

\subsection{A Common Tongue for Uncommon Artefacts}

The final methodological challenge is the hardest to make look easy:
comparing a student project description written by undergraduates with a
peer-reviewed abstract and a patent claims section requires a framework that
abstracts across genre, vocabulary, and register.
A student team describing their ``carbon-sequestration device using engineered
\textit{E.~coli}'' and a patent claiming ``a method for biological conversion
of atmospheric CO\textsubscript{2} via enhanced carbonic anhydrase expression
in recombinant microorganisms'' may be, at some level, describing the same
idea.
No keyword overlap will tell you that.

Neural text embeddings offer a way through this impasse.
\citet{reimers2019sentencebert} demonstrated that sentence-transformer models
produce high-dimensional vector representations in which semantic relatedness
corresponds to geometric proximity.
Documents that mean similar things end up close together in the embedding
space regardless of surface-level vocabulary differences.
The model has learned, from the statistical structure of language across vast
corpora, that certain concepts tend to appear in similar contexts; it does
not need to be told explicitly that ``carbonic anhydrase'' and ``carbon
capture'' belong to the same conceptual neighbourhood.

Projecting these high-dimensional vectors to two dimensions using UMAP
\citep{mcinnes2018umap} and identifying clusters with HDBSCAN
\citep{campello2013hdbscan} produces a semantic map in which artefacts
discussing similar problems cluster together, whether they are student project
descriptions, journal abstracts, or patent claims.
This shared representation is what makes it possible to ask whether the three
artefact types overlap thematically and whether that overlap is geographically
localised.

The present analysis is, then, an attempt to answer a question about the
structure of a nascent field using tools borrowed from machine learning and
applied to a domain they were not originally designed for.
This is, more or less, what all good borrowing looks like.

% ============================================================
\section{Data}
\label{sec:data}
% ============================================================

\subsection{Papers (OpenAlex)}

We retrieved academic papers by searching OpenAlex for synthetic biology
keywords in titles and abstracts. [Describe final corpus size, year range,
geographic coverage.]

\subsection{Patents (Lens.org)}

We retrieved patents from Lens.org using the same keyword list.
[Describe final corpus size, year range, CPC codes.]

\subsection{iGEM projects and parts}

iGEM team projects were downloaded from [source]. Parts were downloaded
from the iGEM Parts Registry. [Describe year range and filtering.]

\subsection{Shared schema}

All artifacts were normalized to a common schema with fields:
\texttt{id}, \texttt{type}, \texttt{title}, \texttt{text}, \texttt{year},
\texttt{city}, \texttt{country}, \texttt{lat}, \texttt{lon},
\texttt{theme\_primary}, \texttt{theme\_secondary},
\texttt{case\_study\_flag}, \texttt{case\_study\_confidence},
\texttt{retrieval\_reason}.

% ============================================================
\section{Methods}
\label{sec:methods}
% ============================================================

\subsection{Embedding}

Each artifact's title and abstract were concatenated and encoded with
\texttt{all-MiniLM-L6-v2}, a sentence-transformer model that produces
384-dimensional dense vectors capturing semantic meaning
\citep{reimers2019sentencebert}.

\subsection{Dimensionality reduction}

Embeddings were projected to 2D with UMAP using the cosine distance metric,
$n\_\text{neighbors} = 15$, and $\text{min\_dist} = 0.1$
\citep{mcinnes2018umap}.

\subsection{Clustering}

Clusters were identified with HDBSCAN \citep{campello2013hdbscan} applied
to the UMAP-reduced coordinates, with $\text{min\_cluster\_size} = 10$.

\subsection{Case study identification}

Artifacts were tagged as belonging to the carbon-capture case study if
their text contained at least one of a predefined keyword list
(carbon capture, carbon sequestration, CO\textsubscript{2} fixation, etc.).

\subsection{Geocoding}

Institution locations were geocoded to city-level latitude/longitude
coordinates using Nominatim (OpenStreetMap).

% ============================================================
\section{Results}
\label{sec:results}
% ============================================================

[To be written after running the pipeline.]

\subsection{Corpus overview}

\subsection{Semantic clustering}

\subsection{City-level patterns}

\subsection{Carbon capture case study}

% ============================================================
\section{Discussion}
\label{sec:discussion}
% ============================================================

[To be written.]

% ============================================================
\section{Limitations}
\label{sec:limitations}
% ============================================================

Several limitations should be noted.
First, keyword-based corpus construction may miss relevant work using
different terminology and may include tangentially relevant work.
Second, geocoding is approximate and some institution locations are missing.
Third, iGEM project descriptions vary in length and quality.
Fourth, we infer semantic relatedness from co-occurrence in embedding space,
not from direct citation links.

% ============================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================

[To be written.]

% ============================================================
\bibliographystyle{apalike}
\bibliography{references}

% ============================================================
\end{document}
